---
title: "Empirical 2"
output: html_document
date: '2022-09-23'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results='hide')
```

```{r, eval = FALSE}
remotes::install_github("traitecoevo/plant", branch = "develop", force = TRUE)
```

```{r load libraries}
library(plant)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(ggrepel)
library(scatterplot3d)
source("R/grow_tidy_expand_totals.R")
source("R/run_mypatch.R")
source("R/thinning_plot.R")
source("R/plot_stand.R")
source("R/find_closest_timestamps.R")
```

```{r activate logger}
plant_log_console()
```

# Compare model output with empirical data

First, grow a stand - traits and other preset values not altered
Initial seed rain is 200, a_l1 = 2.17 (height of plant with LA 1m^2), a_l2 = 0.5 (exponent of relationship between height and LA), lma = 0.1978791, rho = 608, hmat = 16.59587, B_lf1 = 0.8273474, B_lf2 = 0.5, recruitment decay = 5

data is in units m^-2 (density) and m (stem diameter)
```{r}
examplestand <- grow_tidy_expand_totals(a_dG2 = 30)
stand_totals <- examplestand$patch_total  %>%
  mutate(
    study = "plant",
    replicate = 1,
    diameter_stem_av = (diameter_stem_av*100), #convert m to cm
    density=density*10000                      #convert m^2 to ha
   ) %>%
  group_by(study, species, replicate)

stand_totals %>%
  filter(density > 0.01) %>%
  ggplot(aes(diameter_stem_av, density)) +
  geom_line() +
  scale_x_log10(name = "mean stem diameter (cm)") +
  scale_y_log10(name = "density (ha)", labels = scales::comma) +
  theme_classic(base_size = 18)
```

# Generate self-thinning lines from slope and intercept values obtained through statistical models of past papers
data is log transformed and in units ha^-1 (density) and cm (stem diameter)
```{r}
empirical_studies <- read.csv("data/empirical_data.csv", header = TRUE) %>% 
  group_by(study, species, replicate) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>%
  mutate(
    intercept = intercept %>% exp(), #get back un-transformed values, use log10 scale instead of data transformation
    density = as.numeric(intercept)*diameter_stem_av^(slope)
         ) %>%
  filter(study != "brunet_navarro_et_al",
         study != "kurinobu_miyaura")
# remove intercept for kurinobu_miyaura so just use slope and min/max stem diameter vals to plot

empirical_studies %>%
  ggplot(aes(diameter_stem_av, density, group = paste(study, species, replicate))) +
    geom_line(col = "grey") +
    scale_x_log10(labels = scales::comma) +
    scale_y_log10(labels = scales::comma) +
    theme_classic(base_size = 18) +
    xlab("mean stem diameter (cm)") +
    ylab("density (ha)")

#fix_natural_log <- empirical_studies %>% filter(study == "comeau_et_al" | study == "condes_et_al" | study == "aguirre_rio_condes" | study == "forrester_et_al")
```

# Plot self-thinning lines from compilation of datasets of major studies plotted with modelled STL using plant
```{r}
plant_and_empirical <- 
  stand_totals %>% filter(density > 0.01, diameter_stem_av > 0.1) %>%
  ggplot(mapping = aes(diameter_stem_av, density, group = paste(study, species, replicate))) +
  geom_line(data = empirical_studies, col = "grey") +
  geom_line(size = 1, col = "turquoise") +  
  geom_abline(intercept = 1, slope = -3/2, linetype = "dashed") +
  scale_x_log10(labels = scales::comma) +
  scale_y_log10(labels = scales::comma) +
  theme_classic(base_size = 18) +
  xlab("mean stem diameter (cm)") +
  ylab("density (ha)")

ggsave(plant_and_empirical, filename = "../self-thinning/plant_and_empirical.jpeg", width = 6, height = 5)
plot_size_distribution(examplestand$patch_expand$species)

# x and y limits from each study
```


```{r}
plant_and_empirical <- 
  stand_totals %>% filter(density > 0.01, diameter_stem_av > 0.1) %>%
  ggplot(mapping = aes(diameter_stem_av, density, group = paste(study, species, replicate))) +
  geom_line(data = prev_studies, col = "grey") +
  geom_line(size = 1, col = "turquoise") +  
  geom_abline(intercept = 1, slope = -3/2, linetype = "dashed") +
  scale_x_log10(limits=c(0.15,100), labels = scales::comma) +
  scale_y_log10(limits = c(0.001, 100), labels = scales::comma) +
  theme_classic(base_size = 18) +
  xlab("mean stem diameter (cm)") +
  ylab("density (ha)")
```


```{r}
empirical <- read.csv("data/empirical_data.csv", header = TRUE) %>% 
  filter(study != "pretzsch_rio",
         study != "pretzsch_mette",
         study != "brunet_navarro_et_al",
         study != "rivoire_le_moguedec",
         study != "forrester_et_al") %>%
  select(-min_n, -max_n) %>%
  rename("xmin" = "min_dq",
         "xmax" = "max_dq") %>%
  mutate(xmin = log(xmin),
         xmax = log(xmax))
  
# generate list of start and end points for each line
d2 <- empirical %>%
  mutate(id = row_number()) %>%
  mutate(across(starts_with("x"), ~ slope * . + intercept, .names = "y{str_sub(.col, 2, 4)}")) %>%
  pivot_longer(starts_with(c("x", "y"))) %>% 
  separate(name, into = c("var", "position"), sep = 1) %>% 
  pivot_wider(names_from = var)

# data without stem diameter values to add in as abline?
empirical_no_vals <- read.csv("data/empirical_data.csv", header = TRUE) %>%
  subset(., study %in% c("pretzsch_rio", "brunet_navarro_et_al", "rivoire_le_moguedec", "pretzsch_mette")) %>%
  select(-min_dq, - max_dq, -min_n, -max_n) %>%
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>%
  mutate(density = as.numeric(intercept)*diameter_stem_av^(slope))

d2 %>% 
  ggplot(aes(x, y)) +
  geom_line(aes(colour = factor(id))) +
  theme_classic() +
  xlab("log mean stem diameter (cm)") +
  ylab("log density (ha)") +
  theme(legend.position="none")

stand_totals_log_vals <- stand_totals %>% 
  filter(density > 0.01, diameter_stem_av > 0.1) %>%
  mutate(density = log(density),
         diameter_stem_av = log(diameter_stem_av))

d2 %>%
  mutate(
    diameter_stem_av = x,
    density = y
  ) %>%
  ggplot(aes(diameter_stem_av, density)) +
  geom_line(aes(colour = factor(id))) +
  geom_line(data = stand_totals_log_vals) +
  theme_classic() +
  xlab("log mean stem diameter (cm)") +
  ylab("log density (ha)") +
  theme(legend.position="none")
```


didn't work...
```{r}
install.packages("plotrix")
library(plotrix)
#loop through with base r plot?
plot(1:4,1:4,ylim=c(0,4))
abline(intercept, slope)
segments(x0=2,x1=2)

for (i in 1:(length((empirical_studies)-1))) { 
abline(a = empirical_studies$intercept[i], b = empirical_studies$slope[i])
ablineclip(x1 = empirical_studies$min_dq[i], x2 = empirical_studies$max_dq[i])
i <- i + 1
}
```

```{r}
example <- tibble(slope = c(-1.52, -1.42, -1.56, -1.57, -1.57, -1.68, -1.67, -1.6, -1.73, -1.69, -1.79), intercept = c(12.07, 11.29, 12.21, 12.26, 12.14, 12.5, 12.58, 12.28, 12.72, 12.53, 12.29), xmin = c(9.00, 6.00, 6.00, 6.00, 6.00, 9.00, 7.00, 7.00, 7.00, 7.00, 18.00), xmax = c(26.00, 59.00, 53.00, 53.00, 53.00, 63.00, 67.00, 64.00, 64.00, 64.00, 50.00))

example %>%
  ggplot() +
  geom_abline(intercept = example$intercept, slope = example$slope) +
  xlim(example$min_dq, example$max_dq)

plot<-ggplot(empirical_studies,aes(x=empirical_studies$diameter_stem_average,y=empirical_studies$density))
for (i in 1:(length(empirical_studies)-1)) { 
    plot=plot+geom_abline(slope = empirical_studies$slope[i], intercept = empirical_studies$intercept[i]) +
      xlim(empirical_studies$min_dq[i], empirical_studies$max_dq[i])
    i <- i + 1
}

print(plot)
```