---
title: "Plotting empirical data against plant"
output: html_document
date: '2022-08-24'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results='hide')
```

```{r, eval = FALSE}
remotes::install_github("traitecoevo/plant", branch = "develop", force = TRUE)
```

```{r load libraries}
library(plant)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(ggrepel)
library(scatterplot3d)
source("R/grow_tidy_expand_totals.R")
source("R/run_mypatch.R")
source("R/thinning_plot.R")
source("R/plot_stand.R")
source("R/find_closest_timestamps.R")
```

```{r activate logger}
plant_log_console()
```

# Here we compare model output with empirical datasets
First, grow a stand - traits and other preset values not altered
Initial seed rain is 200, a_l1 = 2.17 (height of plant with LA 1m^2), a_l2 = 0.5 (exponent of relationship between height and LA), lma = 0.1978791, rho = 608, hmat = 16.59587, B_lf1 = 0.8273474, B_lf2 = 0.5, recruitment decay = 5
```{r}
examplestand <- grow_tidy_expand_totals()
stand_totals <- examplestand$patch_total  
```

# Generate self-thinning lines from slope and intercept values obtained through statistical models of past papers

Trouve et al stand data with one species' slope and intercept values obtained from 4 different statistical models
Intercept values are logged so take exponential
```{r}
fitted_trouve <- empirical_studies %>% filter(study == "trouve_et_al") %>%
  select(slope, intercept, species, replicate) %>%
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(
    slope = as.numeric(slope),
    intercept = intercept %>% exp(),
    density = as.numeric(intercept)*as.numeric(diameter_stem_av)^as.numeric(slope),
    study = "Trouve_et_al")
```

Condes et al stand data with species slope and intercept values obtained for 2 species (pine, beech) in 6 locations (12 sets of slope values)
Take natural log of intercepts
Max stand denstiy index i.e. max # stems per hectare for reference QMD of 25cm
```{r}
fitted_condes <-
  tibble(
  slope = c(-1.545, -1.539, -2.226, -1.494, -1.592, -1.704, -1.918, -1.469, -2.244, -1.850, -1.831, -1.720),
  intercept = c(190722, 154611, 1213554, 127071, 185631, 210150, 577162, 110800, 1615685, 341925, 428430, 254156), 
  species = c(rep(1, 6), rep(2, 6)),
  replicate = c(1, 2, 3, 4, 5, 6, 1, 2, 3, 4, 5, 6)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Condes_et_al")
```

Weiskittel, Gould & Temesgen stand data
Species slope and intercept values obtained for 3 species (Douglas-fir, Red alder, Western hemlock) using OLS & SFA methods
Density: trees per ha, mean tree size: QMD -- both log transformed
```{r}
fitted_wgt <-
  tibble(
  slope = c(-1.2618, -1.5473, -1.1133, -1.2208, -1.6545, -1.7336),
  intercept = c(10.7509, 11.2087, 9.7315, 10.8665, 12.1299, 13.0391) %>% exp(),
  species = c(1, 1, 2, 2, 3, 3),
  replicate = c(1, 2, 1, 2, 1, 2)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Weiskittel_Gould_Temesgen")
```

Rivoire and Le Moguedec
Oak, beech and spruce modelled with stochastic frontier estimation & generalised
optimisation (-spruce intercept n.a.) (5 STLs)
```{r}
fitted_riv_mog <-
  tibble(
  slope = c(-1.77, -1.76, -1.60, -1.86, -1.59),
  intercept = c(12.22, 11.69, 11.78, 12.17, 11.99) %>% exp(),
  species = c(1, 1, 2, 2, 3),
  replicate = c(1, 2, 1, 2, 1)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Rivoire_Le_Moguedec")
```

Gosper Euc data: E. salubris y = 4.15 - 1.33x
Density per ha, QMD at 10cm above base
```{r}
fitted_gosper <- 
  tibble(
  slope = -1.33,
  intercept = 4.15 %>% exp(), species = seq_len(1)
  ) %>%
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Gosper")
```

Hammond & Niklas
A. alba - use mass not stem diameter...?
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

Eljatib & Weiskittel
```{r}
fitted_elj_weis <-
  tibble(
  slope = c(-1.2741, -1.4742, -1.33),
  intercept = c(11.4787, 12.257, 11.71) %>% exp(),
  species = 1,
  replicate = c(1, 2, 3)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Eljatib_Weiskittel")

# The dynamic - mortality (D-P-NBME) model strategy followed by the dynamic - density model (D-P-LME) gave rather high self-thinning lines. “unsupported by data and current northofagus knowledge” 
# -2.2396, -2.7194
# 14.535, 16.352
```

Pretzsch & Rio 2020
N. spruce, S. pine, D-fir, E. beech, s. oak
Central Europe long-term experimental plots
```{r}
fitted_pret_rio <-
  tibble(
  slope = c(-1.650, -1.789, -1.534, -1.887, -1.753),
  intercept = c(12.560, 12.473, 11.811, 12.865, 12.254) %>% exp(),
  species = seq_len(5)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Pretzsch_Rio")
```

Andrews et. al. 2018
Forests in Acadian Region, Northeastern North America. 15 species of soft and hardwood
```{r}
fitted_andrews <-
  tibble(
  slope = c(-1.96, -1.32, -1.63, -1.57, -1.28, -1.45, -1.38, -1.66, -1.68, -1.77, -1.62, -1.84, -1.39, -1.47, -1.33),
  intercept = c(12.82, 10.98, 11.82, 11.78, 10.98, 11.61, 11.38, 11.88, 11.89, 12.09, 11.70, 12.39, 11.18, 11.57, 10.97) %>% exp(),
  species = seq_len(15)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Andrews_et_al")
```

Aguirre, A., del Río, M., & Condés, S. 2018
Pine species: P. halepensis Mill., P. nigra Arn., P. pinea L., P. pinaster Ait., and P. sylvestris L.
(eqn 1: log(N-max) no climate, eqn 2b: log(N-max) with climate)
```{r}
fitted_ag_rio_cond <-
  tibble(
  slope = c(-1.7262, -1.7950, -2.1220, -2.2647, -1.8292, -1.7009, -1.7943, -1.8759, -1.9825, -1.7298),
  intercept = c(12.607, 12.619, 13.385, 13.426, 12.317, 7.982, 12.642, 11.819, 13.341, 9.784) %>% exp(),
  species = c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5), 
  replicate = c(1, 2, 1, 2, 1, 2, 1, 2, 1, 2)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Aguirre_Río_Condés")
```

Ge, F. et al 2017
China fir (Cunninghamia lanceolata) pure stand plantations, varied site quality
```{r}
fitted_ge <-
  tibble(
  slope = c(-4.298, -3.417, -2.865, -1.590, -1.694),
  intercept = c(18.933, 16.950, 15.695, 12.365, 12.544) %>% exp(),
  species = seq_len(5)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Ge_et_al")
```

Stankova, T. V., & Diéguez-Aranda, U. 2017. Galicia (NW Spain) and Scots pine plantations in Bulgaria.
Even-aged stands of four tree species
Radiata pine (Pinus radiata), Scots pine (Pinus sylvestris), downy birch (Betula pubescens), English oak (Quercus robur)
Use mean stem volume not diameter...
```{r}
fitted_stan_dieg <-
  tibble(
  slope = c(-1.26, -1.03, -1.16, -0.71),
  intercept = c(3621, 6169, 11279, 2321),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Stankova_Diéguez-Aranda")
```

Brunet-Navarro, P., Sterck, F. J., Vayreda, J., Martinez-Vilalta, J., & Mohren, G. M. J. (2016).
The minimum tree density threshold where self-thinning take place was 732 trees ha−1 for P. halepensis, 600 trees ha−1 for P. nigra, 579 trees ha−1 for P. sylvestris and 581 trees ha−1 for 
```{r}
fitted_brunet <-
  tibble(
  slope = c(-0.533, -0.540, -0.589, -0.571),
  intercept = c(14.85, 16.01, 18.31, 18.54) %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Brunet-Navarro_et_al")
```

Vospernik, S., & Sterba, H. 2015
Norway spruce, European beech, Scots pine, European larch, oak, Austrian pine, silver fir, ash, stone pine and hornbeam. Sycamore, grey alder, birch, black locust and black alder are found on less than 100 plots.
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

Wang, N., Gao, J., Xu, S.-S., Zhang, W.-P., & Wang, G.-X. 2013
Mixed montane, evergreen montane, deciduous montane, shrub. Mixed species.
humid area: mean α = -1.74, predicted α =  -1.58
semi-arid area: mean α = -1.27, predicted α =  -1.29
arid area: mean α = -0.95, predicted α = -0.94
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

Charru, M., Seynave, I., Morneau, F., Rivoire, M., & Bontemps, J.-D. 2012
Pure and even-aged plots of 11 broadleaved and coniferous species
```{r}
fitted_charru_et_al <-
  tibble(
  slope = c(-1.809, -1.758, -1.911, -1.941, -1.615, -1.881, -1.653, -1.711, -1.878, -1.779, -1.646),
  intercept = c(12.270, 12.138, 12.681, 12.950, 11.993, 12.512, 12.104, 11.982, 13.086, 12.621, 12.133) %>% exp(),
  species = seq_len(11)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Charru_et_al")
```

Tang, S., Meng, F.-R., & Meng, C. H. 1995
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

Forrester et al 2021
Data from temperate Australian Eucalyptus plantations (436 plots), subtropical forests in China (88 plots), and temperate forests in Switzerland (1055 plots)
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

Morris, Morris & Myerscough data: Ocimum basilicum (herb?)
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

# Combine studies:
```{r}
# correct for units -- cm to m, ha to m^2

prev_studies <- as_tibble(bind_rows(fitted_condes, fitted_trouve, fitted_riv_mog, fitted_wgt, fitted_elj_weis, fitted_ag_rio_cond, fitted_andrews, fitted_charru_et_al, fitted_pret_rio, fitted_stan_dieg)) %>%
  mutate(replicate = replace_na(replicate, 1),
         density=density/10000,
         diameter_stem_av = diameter_stem_av/100)
```

# Typical self thinning line using plant:
```{r}
examplestand <- grow_tidy_expand_totals(hmat = 15, a_dG2 = 30)

patch_totals <- examplestand$patch_total %>%
  mutate(
    study = "plant",
    replicate = 1
    ) %>%
  group_by(study, species)

ggplot(patch_totals, aes(diameter_stem_av, density)) +
  geom_line() +
  scale_x_log10(name = "Mean stem diameter (cm)") +
  scale_y_log10(name = "Density (ha)") +
  theme_classic(base_size = 18)

# plot STL
typical_stand = examplestand$patch_total %>%
  ggplot(aes(diameter_stem_av, density)) +
  geom_line() +
  scale_x_log10(limits = c(0.001, 2.0), labels = scales::comma, name = "log mean stem diameter (m)") +
  scale_y_log10(breaks=c(0.01,1.0,100,1000), limits = c(0.001, 5000), labels = scales::comma, name = "log density " ~ (m^2)) +
  theme_classic(base_size = 18)

ggsave(typical_stand, filename = "../self-thinning/typical_stand.pdf", width = 6, height = 5)
```

# Plot self-thinning lines from compilation of datasets of major studies plotted with modelled STL using plant
```{r}
plant_and_empirical = ggplot(mapping = aes(diameter_stem_av, density, group = paste(study, species, replicate))) +
  geom_line(data = prev_studies, col = "grey") +
  geom_line(data = patch_totals, size = 1, col = "green") + 
  geom_abline(intercept = log(0.085), slope = -3.7/2, linetype = "dashed") +
  scale_x_log10(limits = c(0.001, 2.0), labels = scales::comma, name = "log mean stem diameter (m)") +
  scale_y_log10(breaks=c(0.01,1.0,100,1000), limits = c(0.001, 5000), labels = scales::comma, name = "log density " ~ (m^2)) +
  theme_classic(base_size = 18)


ggsave(plant_and_empirical, filename = "../self-thinning/plant_and_empirical.jpeg", width = 6, height = 5)
plot_size_distribution(examplestand$patch_expand$species)

# x and y limits from each study
```

```{r}
ggplot(mapping = aes(diameter_stem_av, density)) +
  geom_abline(intercept = 1, slope = -3/2, linetype = "dashed") +
  geom_line(data = prev_studies) +
  scale_x_log10(limits = c(0.1, 500), labels = scales::comma) +
  scale_y_log10(limits = c(0.1, 500), labels = scales::comma) +
  theme_classic(base_size = 18) +
  xlab("mean stem diameter (cm)") +
  ylab("density (ha)")

```


?? fitted_brunet 
intercept = c(14.85, 16.01, 18.31, 18.54) %>% exp()

Brunet-Navarro, P., Sterck, F. J., Vayreda, J., Martinez-Vilalta, J., & Mohren, G. M. J. (2016). Self-thinning in four pine species: An evaluation of potential climate impacts. Annals of Forest Science, 73(4), 1025–1034. 
https://doi.org/10.1007/s13595-016-0585-y
https://annforsci.biomedcentral.com/articles/10.1007/s13595-016-0585-y/tables/2

?? fitted_ge
https://www.mdpi.com/1999-4907/8/10/355/htm

?? fitted_gosper
intercept = 4.15 %>% exp()

Gosper, C. R., Yates, C. J., Cook, G. D., Harvey, J. M., Liedloff, A. C., McCaw, W. L., Thiele, K. R., & Prober, S. M. (2018). A conceptual model of vegetation dynamics for the unique obligate-seeder eucalypt woodlands of south-western Australia. Austral Ecology, 43(6), 681–695. https://doi.org/10.1111/aec.12613
