---
title: "Self-thinning lines"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=FALSE}
#remotes::install_github("traitecoevo/plant", branch = "develop", force = TRUE)
```

```{r load libraries}
library(plant)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(purrr)
source("R/grow_tidy_expand_totals function.R")
source("R/base_params function.R")
source("R/run_mypatch_edit function.R")
source("R/grow_plot function.R")
```

```{r activate logger}
plant_log_console()
```


## 1. Grow stand: high resolution cohort spacing + recruitment decay = 5
```{r}
allsteps <- grow_tidy_expand_totals(oderel=1e-4, odeabs=1e-4, scheps=0.0005, rec_decay=5)

allsteps$patch_total %>%
  ggplot(aes(time, individuals)) +
  geom_line()

allsteps$patch_total %>%
  ggplot(aes(time, area_leaf_av)) +
  geom_line()

```


## 2. Self-thinnning Plot
Average size over time. Get average size by dividing total of chosen size variable by the sum of individuals per unit area - we have mean stem area, mean stem diameter, mean height, mean mass above ground, and also mean LA

```{r}
# self-thinning line:
allsteps$patch_total  %>% 
  filter(time > 0.2) %>%
  ggplot(aes(area_stem_av, individuals, col=time)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

allsteps$patch_total  %>% 
  filter(time > 0.2) %>%
  ggplot(aes(height_av, individuals, col=time)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

# average stem area over time
allsteps$patch_total  %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, area_stem_av)) +
  geom_line() +
  scale_y_log10() +
  theme_classic()
 
allsteps$patch_expand$species %>%
  plot_size_distribution() +
  geom_line()

# first wave of individuals captured
allsteps$patch_total %>% filter(height > 0, time> 0, time<150) %>%
  ggplot(aes(time, individuals, colour=height)) +
  geom_line() + 
  theme_classic()

```


## Varying mortality rate

```{r}
mort_runs <- c(0.0, 0.1) %>% set_names(.) %>% map(~grow_tidy_expand_totals(B_dI1 = .x))

mort_values <-
  mort_runs %>% map_dfr(~.x$patch_total, .id = "run")
  
mort_values %>%
  ggplot(aes(time, area_leaf, col=run)) +
  geom_line()


```


## Trait comparison: 1. LMA

```{r}

# 1. thick vs thin leaves
lma_runs <- 
    c(0.07, 0.24, 0.5) %>% set_names(.) %>% 
    map(~grow_tidy_expand_totals(traits=trait_matrix(.x, "lma")))

lma_values <-
  lma_runs %>% map_dfr(~.x$patch_total, .id = "run")

lma_values %>%
  filter(time > 0.2) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

lma_values %>%
  filter(time > 0.2, area_leaf_av > 1, area_leaf_av < 100) %>%
  ggplot(aes(area_leaf_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

#density over time both stands
lma_values %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, area_leaf, col=run)) + 
  geom_line() +
  theme_classic()


#lma_runs[["patch_total"]], lma_runs$"0.07"$patch_total   =pull out stuff
```


## Trait comparison: 2. narea ...also blf1

```{r}

narea_runs <- c(0.00187, 0.003, 0.005) %>% set_names(.) %>% map(~grow_tidy_expand_totals(narea = .x))
narea_vals <- narea_runs %>% map_dfr(~.x$patch_total, .id = "run")

narea_vals %>%
  filter(time > 0.2) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()


blf1_runs <- c(0.4, 0.6, 1) %>% set_names(.) %>% map(~grow_tidy_expand_totals(B_lf1 = .x))
blf1_vals <- blf1_runs %>% map_dfr(~.x$patch_total, .id = "run")

blf1_vals %>%
  filter(time > 0.2) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

```


## Trait comparison: 3. HMAT

```{r}

hmat_runs <- c(5, 10, 25, 40) %>% set_names(.) %>% map(~grow_tidy_expand_totals(traits=trait_matrix(.x, "hmat")))
hmat_vals <-
  hmat_runs %>% map_dfr(~.x$patch_total, .id = "run")

hmat_vals %>%
  filter(time > 0.2) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()
```



### TASKS ###

1. plot total la and biomass over time - differences? how big?
2. turn into differences in self-thinning line, zoom in and out
3. vary multiple traits at once
4. before zoomed out log-log plot check differences are there or not
5. check stand:
   a. with and without recruitment decay 
   b. check LA and biomass over time, LA should be bigger than 0.5



# 1. plot total la and biomass over time - differences? how big?

```{r}

lma_runs <- c(0.07, 0.24, 0.5) %>% set_names(.) %>% map(~grow_tidy_expand_totals(traits=trait_matrix(.x, "lma")))
lma_values <-
  lma_runs %>% map_dfr(~.x$patch_total, .id = "run")

#LA over time
lma_values %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, area_leaf, col=run)) + 
  geom_line() +
  theme_classic()

#total mass over time
lma_values %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, mass_total, col = run)) +
  geom_line() +
  theme_classic()

#average mass over time
lma_values %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, mass_above_ground_av, col = run)) +
  geom_line() +
  theme_classic()

lma_values %>%
  filter(time > 0.2, time < 15) %>%
  ggplot(aes(time, mass_above_ground_av, col = run)) +
  geom_line() +
  theme_classic()



narea_runs <- c(0.00187, 0.003, 0.005) %>% set_names(.) %>% map(~grow_tidy_expand_totals(narea = .x))
narea_vals <- narea_runs %>% map_dfr(~.x$patch_total, .id = "run")

# LA over time
narea_vals %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, area_leaf, col=run)) + 
  geom_line() +
  theme_classic()

# total mass over time
narea_vals %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, mass_total, col = run)) +
  geom_line() +
  theme_classic()

# average mass over time
narea_vals %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, mass_av, col = run)) +
  geom_line() +
  theme_classic()


# plot size distribution
lma_vals_expand <- lma_runs %>% map_dfr(~.x$patch_expand$species, .id = "run")
plot_size_distribution(lma_vals_expand)

narea_vals_expand <- narea_runs %>% map_dfr(~.x$patch_expand$species, .id = "run")
plot_size_distribution(narea_vals_expand)

```


# 2. turn into differences in self-thinning line, zoom in and out
```{r}
# self-thinning plots
lma_values %>%
filter(time > 0.2) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

# zoom in/out
lma_values %>%
  mutate(area_leaf_av = area_leaf/individuals) %>%
filter(time > 0.2, area_leaf_av > 1, area_leaf_av < 100) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

```


# 3. vary two traits at once: LMA/narea
```{r}
# 3.a. vary 2 TRAITS: LMA + narea

lmas <- c(0.07, 0.24, 0.5)
nareas <- c(0.00187, 0.003, 0.005)   #x and y equal length, or one vector of length 1 will get repeated

lmas_nareas <- map2(lmas, nareas, ~grow_tidy_expand_totals(traits=trait_matrix(.x, "lma"), narea = .y), .id = "run")

lmas_nareas_totals <- lmas_hmats %>% map_dfr(~.x$patch_total, .id = "run") 
## anonymous function takes .x from before pipe and feeds into $patch_total, .x is the lmas_hmats object, map_dfr needs a function so can't just do map_dfr(lmas_hmats$patch_total)
lmas_nareas_expandsp <- lmas_hmats %>% map_dfr(~.x$patch_expand$species, .id = "run")

lmas_nareas_totals %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, area_leaf, col=run)) + 
  geom_line() +
  theme_classic()

lmas_nareas_totals %>%
  filter(time > 0.2) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

lmas_nareas_totals %>%
  filter(time > 0.2, area_stem_av > 1e-06, area_stem_av < 1e0) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

p1 = lmas_nareas_expandsp %>% filter(run == 1) %>% plot_size_distribution()

```


# 3.b. vary 3 traits: LMA/narea/hmat
```{r}

lmas2 <- c(0.07, 0.125, 0.24, 0.5)
hmats2 <- c(5, 10, 25, 40)
nareas2 <- c(0.00187, 0.0025, 0.0035, 0.005)

tradeoffs <- list(c(0.07, 0.125, 0.24, 0.5), c(5, 10, 25, 40), c(0.00187, 0.0025, 0.0035, 0.005))
tradeoffs_runs <- pmap(tradeoffs, ~grow_tidy_expand_totals(traits=trait_matrix(..1, "lma"), trait_matrix(..2, "hmat"), narea = ..3))

tradeoffs_values <- tradeoffs_runs %>% map_dfr(~.x$patch_total, .id = "run")

tradeoffs_values %>%
  filter(time > 0.2, area_stem_av >1e-07, area_stem_av < 1e-00) %>%
  ggplot(aes(area_stem_av, individuals, col=run)) + 
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()


tradeoffs_values %>%
  filter(time > 0.2) %>%
  ggplot(aes(time, area_leaf, col=run)) + 
  geom_line() +
  theme_classic()

```


#use code in make_hyperpar to plot tradeoffs
#e.g. LMA vs k_l
#narea affecting Amax and respiration
#wood density affecting DI mortality and wood turnover rate
# -> vary strength of tradeoffs and plot stronger vs weaker tradeoff strength
# exponent = 0 ?

#run hyperparameter function to see what changes, look into the function, lma, narea and wood density changes stuff, parameter for strength of tradeoff 


lma/ll relationship:
k_l   <- B_kl1 * (lma / lma_0) ^ (-B_kl2) = rate of leaf turnover * (LMA/mean LMA) ^ scaling exponent
lma_0: central (mean) value for LMA
B_kl1: rate of leaf turnover at lma_0
B_kl2: scaling slope(exponent) for phi (LMA) in leaf turnover = 1.71 default
# if lma = lma_0 or scaling exponent = 0, k_l = B_kl1

WD/mortality relationship:
d_I  <- B_dI1 * (rho / rho_0) ^ (-B_dI2) = rate of inst. mortality * (rho/mean rho) ^ scaling exponent
rho_0: central (mean) value for WD
B_dI1: rate of instantaneous mortality at rho_0
B_dI2: scaling slope for WD in intrinsic mortality = 0 default
# if rho = rho_0 or scaling exponent = 0, d_I = B_dI1

WD/wood turnover relationship:
k_s  <- B_ks1 *  (rho / rho_0) ^ (-B_ks2) = rate of sapwood turnover * (rho/mean rho) ^ scaling exponent
B_ks1: rate of sapwood turnover at rho_0
B_ks2: scaling slope(exponent) for rho in sapwood turnover = 0 default
# if rho = rho_0 or scaling exponent = 0, k_s = B_ks1

Narea, photosynthesis, respiration relationship:
Amax <- B_lf1 * (narea/narea_0) ^  B_lf5 = 
narea: nitrogen per leaf area
narea_0: central (mean) value for nitrogen per leaf area
B_lf1: potential CO2 photosynthesis at average leaf N
B_lf5: scaling exponent for leaf nitrogen in maximum leaf photosynthesis = 1 default
# if narea = narea_0 or scaling exponent = 0, Amax = B_lf1

{m} A matrix of trait values
{s} A strategy object
{k_I} light extinction coefficient

## default values of interest:
make_FF16_hyperpar(
    lma_0=0.1978791, 
    B_kl1=0.4565855, 
    B_kl2=1.71, 
    rho_0=608.0, 
    B_dI1=0.01, 
    B_dI2=0.0,
    B_ks1=0.2,
    B_ks2=0.0,
    narea=1.87e-3,
    narea_0=1.87e-3,
    B_lf1=5120.738 * 1.87e-3 * 24 * 3600 / 1e+06,
    B_lf5=1)

#use code in make_FF16_hyperpar to plot tradeoffs
#LMA vs k_l
#narea affecting Amax and respiration
#wood density affecting DI mortality and wood turnover rate

# -> vary strength of tradeoffs and plot stronger vs weaker tradeoff strength

#run hyperparameter function to see what changes, look into the function, lma, narea and wood density changes stuff, parameter for strength of tradeoff 

```{r}

## LMA vs k_l

k_l_tradeoffs <- function(lma_0 = 0.1978791,
                lma = 0.1978791,
                B_kl1=0.4565855,
                B_kl2=1.71){
  
  k_l_val <- B_kl1 * (lma / lma_0) ^ (-B_kl2)
  k_l_df <- tibble(lma, k_l_val, B_kl2)
  return(k_l_df)
  
}

## a. weaker trade-off strength

lmall_tradeoff_low <- seq(0.07, 2.5, by=0.02) %>% set_names(.) %>% map_dfr(~k_l_tradeoffs(lma=.x, B_kl2 = 0.5)) %>% ggplot(aes(lma, k_l_val)) + 
  geom_line() +
  theme_classic()

## b. higher trade-off strength

lmall_tradeoff_high <- seq(0.07, 2.5, by=0.02) %>% set_names(.) %>% map_dfr(~k_l_tradeoffs(lma=.x, B_kl2=2)) %>% ggplot(aes(lma, k_l_val)) + 
  geom_line() +
  theme_classic() 

## narea affecting Amax and respiration

amax_tradeoffs <- function(narea_0=1.87e-3, 
                 narea=1.87e-3, 
                 B_lf5=1,
                 B_lf1=0.8273474){
  
Amax <- B_lf1 * (narea/narea_0) ^ B_lf5
Amax_df <- tibble(Amax, narea, B_lf5)
return(Amax_df)

}

## a. weaker tradeoff strength:
amax_tradeoff_low <- seq(0.00187, 0.0031, by=0.0001) %>% set_names(.) %>% map_dfr(~amax_tradeoffs(narea=.x)) %>% ggplot(aes(narea, Amax)) + 
  geom_line() +
  theme_classic()

## b. higher tradeoff strength:
amax_tradeoff_high <- seq(0.00187, 0.0031, by=0.0001) %>% set_names(.) %>% map_dfr(~amax_tradeoffs(narea=.x, B_lf5=5)) %>% ggplot(aes(narea, Amax)) + 
  geom_line() +
  theme_classic()


## wood density affecting DI mortality and wood turnover rate

wd_tradeoffs <- function(
  rho_0 = 608,
  rho = 608, 
  B_dI1 = 0.01,
  B_dI2=0.0,
  B_ks1=0.2,
  B_ks2=0.0
) {
  
  d_I <- B_dI1 * (rho / rho_0) ^ (-B_dI2)
  k_s <- B_ks1 * (rho / rho_0) ^ (-B_ks2)
  
  di_ks_df <- tibble(k_s, d_I, rho, B_ks2, B_dI2)
  
  return(di_ks_df)
  
}

## a. weaker tradeoff strength

dI_weak <- seq(400, 800, by=10) %>% set_names(.) %>% map_dfr(~wd_tradeoffs(rho=.x, B_dI2=0.5)) %>% ggplot(aes(rho, d_I)) + 
  geom_line() +
  theme_classic()

ks_weak <- seq(400, 800, by=10) %>% set_names(.) %>% map_dfr(~wd_tradeoffs(rho=.x, B_ks2=0.5)) %>% ggplot(aes(rho, k_s)) + 
  geom_line() +
  theme_classic()

## b. higher tradeoff strength

dI_strong <- seq(400, 800, by=10) %>% set_names(.) %>% map_dfr(~wd_tradeoffs(rho=.x, B_dI2=3)) %>% ggplot(aes(rho, d_I)) + 
  geom_line() +
  theme_classic()

ks_strong <- seq(400, 800, by=10) %>% set_names(.) %>% map_dfr(~wd_tradeoffs(rho=.x, B_ks2=3)) %>% ggplot(aes(rho, k_s)) + 
  geom_line() +
  theme_classic()

```


leaf area declines over time.
1. as you get taller, your shade tolerance decreases (see Falster et al 2018 "how functional traits..."). 
So if Hmat is high, leaf area keeps decreasing,
background mortality keeps slowly killing plants, so varying background mortality kills things more/less

higher costs, carbon budget decrease, less shade tolerance
min light level getting higher
thinning budgets 

pushing plants to zero carbon budget the higher HMAT

resource or light compensation point as function of height - solve for this


# 4.before zoomed out log-log plot check differences are there or not
```{r}



```

# 5.check stand:
    a. with and without recruitment decay 
    b. check LA and biomass over time, LA should be bigger than 0.5
```{r}
### 1. with recruitment decay

with_recdecay <- grow_tidy_expand_totals(oderel=1e-4, odeabs=1e-4, scheps=0.0005, rec_decay=5)

with_recdecay$patch_total %>%
  ggplot(aes(time, individuals)) +
  geom_line()

with_recdecay$patch_total %>%
  ggplot(aes(time, area_leaf_av)) +
  geom_line()

lma_norec <- c(0.07, 0.24, 0.5) %>% map(~grow_tidy_expand_totals(traits=trait_matrix(.x, "lma"), rec_deay = 5))
lma_vals_norec <- lma_norec %>% map_dfr(~.x$patch_total, .id = "run")

narea_norec <- c(0.00187, 0.003, 0.005) %>% map(~grow_tidy_expand_totals(narea = .x, , rec_decay = 5))
narea_vals_norec <- narea_norec %>% map_dfr(~.x$patch_total, .id = "run")

### 2. without recruitment decay

## LMAs 
lma_norec <- c(0.07, 0.24, 0.5) %>% map(~grow_tidy_expand_totals(traits=trait_matrix(.x, "lma"), rec_decay = 0))
lma_vals_norec <- lma_norec %>% map_dfr(~.x$patch_total, .id = "run")

with_recdecay$patch_total %>%
  ggplot(aes(time, individuals)) +
  geom_line()

with_recdecay$patch_total %>%
  ggplot(aes(time, area_leaf_av)) +
  geom_line()

## NAREAs
narea_norec <- c(0.00187, 0.003, 0.005) %>% map(~grow_tidy_expand_totals(narea = .x, rec_decay = 0))
narea_vals_norec <- narea_norec %>% map_dfr(~.x$patch_total, .id = "run")

narea_vals_norec %>%
  ggplot(aes(time, individuals, col=run)) +
  geom_line()

narea_vals_norec %>%
  ggplot(aes(time, area_leaf, col=run)) +
  geom_line()

```



Traits again...

## STRATEGIES

## how to implement some strategies? Can I call on the hyperparameter function??

lma affects a few places in the model; see the source code, but only really for converting from leaf area to leaf mass (leaf mass being leaf area multiplied by leaf mass per unit leaf area). However, as a component of the leaf economic spectrum we imagine LMA as affecting a number of other components of the model.

We capture this through what we call a “hyper-parameterisation”; additional parameters and functions that mean that changing one parameter might affect a number of other lower-level parameters. Our default hyper-parameterisation is via the FF16_hyperpar function:
  FF16_hyperpar

  FF16_hyperpar(trait_matrix(0.1, "lma"), FF16_Strategy())

You will rarely need to call this function directly (see below) but note how setting of lma affects parameters k_l, a_p1 and r_l
These are: * k_l: Turnover rate for leaves * a_p1: Leaf photosynthesis per area * r_l: Leaf respiration per mass

This means that it is possible to implement different trade-offs between parameters relatively easily by modifying the hyper-parameterisation function in R, rather than having to modify the underlying physiological model in C++.

To make use of the hyper-parameterisation, the preferred way of setting parameters is through the utility functions strategy and strategy_list. These take a Parameters object:

  p <- FF16_Parameters()
  identical(p$hyperpar, FF16_hyperpar)
  class(p$strategy_default)

This (FF16_Strategy) is the Strategy object that all others will be built from by difference. Running:

  s <- strategy(trait_matrix(0.1, "lma"), p)
  
will create a strategy s where lma is set but also all the parameters that depend on lma.

  lma <- trait_matrix(seq(0.1, 0.5, length.out=5), "lma")

  FF16_hyperpar(trait_matrix(lma, "lma"), s)


```{r}

lmas_hyperpar <- c(0.07, 0.24, 0.4) %>% map_df(~grow_tidy_expand_totals(patch = run_mypatch_edit(hyper_par_fn = make_FF16_hyperpar(lma_0 = .x))))

hyp <- c(0.07, 0.24, 0.4) %>% make_FF16_hyperpar(lma_0=.x)



lmas_hyperpar <- c(0.07, 0.24, 0.4) %>% map(~grow_tidy_expand_totals(patch = run_mypatch_edit(hyper_par_fn = FF16_hyperpar(trait_matrix(0.1, "lma"), s))))

FF16_hyperpar(trait_matrix(0.1, "lma"), strat)
# FF16_hyperpar(m, s, filter = TRUE)


```



## Paired traits
```{r}



```


## Slope, Intercept, etc, values...
```{r}



```

#lma_runs[["patch_total"]]  #pull out stuff