---
title: "Model assessment"
output: html_document
date: '2022-07-28'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results='hide')
```

```{r, eval = FALSE}
remotes::install_github("traitecoevo/plant", branch = "develop", force = TRUE)
```

```{r load libraries}
library(plant)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(ggrepel)
library(scatterplot3d)
source("../self-thinning/R/grow_tidy_expand_totals")
source("../self-thinning/R/base_params function.R")
source("../self-thinning/R/run_mypatch.R")
source("../self-thinning/R/grow_plot function.R")
source("../self-thinning/R/thinning_plot.R")
source("../self-thinning/R/plot_stand.R")
source("../self-thinning/R/find_closest_timestamps.R")
```

```{r activate logger}
plant_log_console()
```


# Here we will systematically analyse and test all aspects of model implementation (varied stand permutations) and also compare with real-world datasets

## First, grow a stand - relatively unaltered in terms of varifying trait and other preset values

## Initial seed rain is 200, a_l1 = 2.17 (height of plant with LA 1m^2), a_l2 = 0.5 (exponent of relationship between height and LA), lma = 0.1978791, rho = 608, hmat = 16.59587, recruitment decay = 5, ...
```{r}
examplestand <- grow_tidy_expand_totals(oderel=1e-4, odeabs=1e-4, scheps=0.001)

# We removed integration part of grow_tidy_expand_totals function, checked integrations match up:
# Patch 1 using integrate_over_size_distribution() function, Patch 2 using plant:::trapezium function directly: 

patch_total1 <- 
  examplestand$patch_expand$species %>% 
  integrate_over_size_distribution() %>%
    mutate(
      # create average sizes
      across(c(diameter_stem, area_stem, area_leaf, height, mass_above_ground), ~.x/density, 
      .names = "{.col}_av")
    )

patch_total2 <- 
  examplestand$patch_expand$species %>% 
  na.omit() %>% 
  filter(step>1) %>% 
  group_by(step, species) %>% 
  summarise(
    density2 = -plant:::trapezium(height, density),
    diameter_stem_av = -plant:::trapezium(height, density * diameter_stem)/ density2
  )%>% ungroup()

bind_cols(patch_total1 %>% select(density, diameter_stem_av), patch_total2)

patch_total1 %>%
  ggplot(aes(time, density)) +
  geom_line()

patch_total1 %>%
  ggplot(aes(time, area_leaf)) +
  geom_line()

## both stands give same totals output
```


# Create some data using trouve et al stand data with one species' slope and intercept values obtained from 4 different statistical models
## intercept values are logged so take exponential
```{r}
fitted_trouve <-
  tibble(
  slope = c(-1.52, -1.42,-1.56, -1.57),
  intercept = c(12.07, 11.29, 12.21, 12.26) %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "trouve_et_al")
```

# Data using Condes et al stand data with species slope and intercept values obtained for 2 species (pine, beech) in 6 locations (12 sets of slope values)
## Take natural log of intercepts
## Max stand denstiy index i.e. max # stems per hectare for reference QMD of 25cm
```{r}
fitted_condes <-
  tibble(
  slope = c(-1.545, -1.539, -2.226, -1.494, -1.592, -1.704, -1.918, -1.469, -2.244, -1.850, -1.831, -1.720),
  intercept = c(190722, 154611, 1213554, 127071, 185631, 210150, 577162, 110800, 1615685, 341925, 428430, 254156) %>% log(), species = c(rep(1, 6), rep(2, 6))
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "condes_et_al")
```

# Data using Weiskittel, Gould & Temesgen stand data with species slope and intercept values obtained for 3 species (Douglas-fir, Red alder, Western hemlock) using OLS & SFA methods
## density: trees per ha, mean tree size: QMD -- both log transformed
```{r}
fitted_wgt <-
  tibble(
  slope = c(-1.2618, -1.5473, -1.1133, -1.2208, -1.6545, -1.7336),
  intercept = c(10.7509, 11.2087, 9.7315, 10.8665, 12.1299, 13.0391) %>% exp(),
  species = c(1, 1, 2, 2, 3, 3)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "weisk_gould_temes")
```

# Data using Rivoire and Le Moguedec
## Oak, beech and spruce modelled with stochastic frontier estimation & generalised
## optimisation (-spruce intercept n.a.) (5 STLs)
```{r}
fitted_riv_mog <-
  tibble(
  slope = c(-1.77, -1.76, -1.60, -1.86, -1.59),
  intercept = c(12.22, 11.69, 11.78, 12.17, 11.99) %>% exp(),
  species = c(1, 1, 2, 2, 3)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "rivoire_moguedec")
```

# Gosper Euc data: E. salubris y = 4.15 - 1.33x
## density per ha, QMD at 10cm above base
```{r}
fitted_gosper <- 
  tibble(
  slope = -1.33,
  intercept = 4.15 %>% exp(), species = seq_len(1)
  ) %>%
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "gosper")
```

# More data?
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

# Plot self-thinning lines from compilation of datasets of major studies plotted with modelled STL
## grouped by study, want study + species
```{r}
examplestand2 <- grow_tidy_expand_totals(oderel=1e-4, odeabs=1e-4, scheps=0.001)

prev_studies <- as_tibble(bind_rows(fitted_condes, fitted_gosper, fitted_trouve, fitted_riv_mog, fitted_wgt))

examplestand2$patch_total %>%
  mutate(
    diameter = diameter_stem_av*100, 
    density=density*10000,
    study = "plant_model"
    ) %>% 
  ggplot(aes(diameter_stem_av, density, group = study, col = study)) +
  geom_line() + 
  geom_abline(intercept = 1, slope = -3/2, linetype = "dashed") +
  geom_line(data = prev_studies) +
  scale_x_log10(limits = c(1e-03, 1e03)) +
  scale_y_log10() +
  theme_classic(base_size = 18) +
  ggtitle("Self-thinning plot") +
  xlab("Average Stem Diameter (m)") +
  ylab("Density")

plot_size_distribution(examplestand2$patch_expand$species)
```


# Check key measures over time for each stand - total LA, biomass, heights, density
## DI mortality
```{r}
plot_stand(dI_total, xvar = "time", yvar = "density", title="Total individuals over time")

plot_stand(dI_total, xvar = "time", yvar = "area_leaf", title="Total Leaf Area over time")

plot_stand(dI_total, xvar = "time", yvar = "height_av", title="Average height over time")

plot_stand(dI_total, xvar = "time", yvar = "mass_total", title="Total mass over time")

plot_stand(dI_total, xvar = "time", yvar = "diameter_stem_av", title="Average stem diameter over time")

```


# We want to test the effect of varying below model parameters on self-thinning lines

# Initial seed rain is 200, a_l1 = 2.17 (height of plant with LA 1m^2), a_l2 = 0.5, k_l = 0.5, lma = 0.1978791, rho = 608, hmat = 16.59587, recruitment decay = 5, ...

## a_l1: Height of plant with leaf area of 1m^2 - default = 2.17
## a_l2: Exponent of relationship between height and leaf area - default = 0.5
## birth rate: default = 200
## k_l: Turnover rate for leaves per year - default = 0.4565855
## b_lf1: Potential CO2 photosynthesis at average leaf nitrogen, ν0, mol per day
##        per m^2 - default = 0.8273474
## b_lf2: Curvature of light response curve - default = 0.5
## d_I: Intrinsic or growth-independent mortality per year - default = 0.01


## Checking values with a_l1
### 2, 3, 5
```{r}
al1_runs <- c(2, 3, 5) %>% set_names(.) %>% map(~grow_tidy_expand_totals(a_l1 = .x))

al1_total <-
  al1_runs %>% map_dfr(~.x$patch_total, .id = "run") %>%
  mutate(run = recode(run, '1' = 'a_l1 = 2', '2' = 'a_l1 = 3', '3' = 'a_l1 = 5'))

##################### ##################### #####################
  p0_1 <- scm_base_parameters("FF16")
  p0_1$strategy_default$a_l1 <- 1
  p0_1$strategy_default$a_l2 <- 0.5
  p0_1$strategy_default$recruitment_decay <- 5
  
  control = scm_base_control()
  control$ode_tol_rel <- 1e-4
  control$ode_tol_abs <- 1e-4
  control$schedule_eps <- 0.0005
  
  patch1 <- run_mypatch_edit(p0 = p0_1, ctrl = control) %>%
    tidy_patch() %>%
    FF16_expand_state()
  
  patch1_total <- 
    patch1$species %>% 
    integrate_over_size_distribution() %>%
    mutate(
        diameter_stem = (2 * sqrt (area_stem / pi)),
        diameter_stem_av = diameter_stem/density,
        across(c(diameter_stem, area_stem, area_leaf, height, mass_above_ground), 
        ~.x/density, .names = "{.col}_av")
    )

p1 = patch1_total %>%
  filter(time > 2) %>%
  ggplot(aes(diameter_stem_av, density)) + 
  geom_line() + 
  scale_x_log10() +
  scale_y_log10() +
  theme_classic(base_size = 16) +
  ggtitle("a_l1 (height of plant with LA of 1m^2) = 2m")
  
##################### ##################### #####################
  p0_2 <- scm_base_parameters("FF16")
  p0_2$strategy_default$a_l1 <- 3
  p0_2$strategy_default$a_l2 <- 0.5
  p0_2$strategy_default$recruitment_decay <- 5
  
  control = scm_base_control()
  control$ode_tol_rel <- 1e-4
  control$ode_tol_abs <- 1e-4
  control$schedule_eps <- 0.0005
  
  patch2 <- run_mypatch_edit(p0 = p0_2, ctrl = control) %>%
    tidy_patch() %>%
    FF16_expand_state()
  
  patch2_total <- 
    patch2$species %>% 
    integrate_over_size_distribution() %>%
    mutate(
        diameter_stem = (2 * sqrt (area_stem / pi)),
        diameter_stem_av = diameter_stem/density,
        across(c(diameter_stem, area_stem, area_leaf, height, mass_above_ground), 
        ~.x/density, .names = "{.col}_av")
    )  
  
  p2 = patch2_total %>%
    filter(time > 2) %>%
    ggplot(aes(diameter_stem_av, density)) + 
    geom_line() + 
    scale_x_log10() +
    scale_y_log10() +
    theme_classic(base_size = 16) +
    ggtitle("a_l1 (height of plant with LA of 1m^2) = 3m")
  
##################### ##################### #####################
  p0_3 <- scm_base_parameters("FF16")
  p0_3$strategy_default$a_l1 <- 5
  p0_3$strategy_default$a_l2 <- 0.5
  p0_3$strategy_default$recruitment_decay <- 5
  
  control = scm_base_control()
  control$ode_tol_rel <- 1e-4
  control$ode_tol_abs <- 1e-4
  control$schedule_eps <- 0.0005
  
  patch3 <- run_mypatch_edit(p0 = p0_3, ctrl = control) %>%
    tidy_patch() %>%
    FF16_expand_state()
  
  patch3_total <- 
    patch3$species %>% 
    integrate_over_size_distribution() %>%
    mutate(
        diameter_stem = (2 * sqrt (area_stem / pi)),
        diameter_stem_av = diameter_stem/density,
        across(c(diameter_stem, area_stem, area_leaf, height, mass_above_ground), 
        ~.x/density, .names = "{.col}_av")
    ) 

p3 = patch3_total %>%
    filter(time > 2) %>%
    ggplot(aes(diameter_stem_av, density)) + 
    geom_line() + 
    scale_x_log10() +
    scale_y_log10() +
    theme_classic(base_size = 16) +
    ggtitle("a_l1 (height of plant with LA of 1m^2) = 5m")
```

## Checking values with a_l2
### 0.3, 0.5, 0.8
```{r}
al2_runs <- c(0.3, 0.5, 0.8) %>% set_names(.) %>% map(~grow_tidy_expand_totals(p = base_params(a_l2 = .x)))

al2_total <-
  al2_runs %>% map_dfr(~.x$patch_total, .id = "run") %>%
  mutate(run = recode(run, '1' = 'a_l2 = 0.3', '2' = 'a_l2 = 0.5', '3' = 'a_l2 = 0.8'))

##################### ##################### #####################
  p0_11 <- scm_base_parameters("FF16")
  p0_11$strategy_default$a_l1 <- 2.17
  p0_11$strategy_default$a_l2 <- 0.3
  p0_11$strategy_default$recruitment_decay <- 5
  
  control = scm_base_control()
  control$ode_tol_rel <- 1e-4
  control$ode_tol_abs <- 1e-4
  control$schedule_eps <- 0.0005
  
  patch11 <- run_mypatch_edit(p0 = p0_11, ctrl = control) %>%
    tidy_patch() %>%
    FF16_expand_state()
  
  patch11_total <- 
    patch11$species %>% 
    integrate_over_size_distribution() %>%
    mutate(
        diameter_stem = (2 * sqrt (area_stem / pi)),
        diameter_stem_av = diameter_stem/density,
        across(c(diameter_stem, area_stem, area_leaf, height, mass_above_ground), 
        ~.x/density, .names = "{.col}_av")
    )

p11 = patch11_total %>%
  filter(time > 2) %>%
  ggplot(aes(diameter_stem_av, density)) + 
  geom_line() + 
  scale_x_log10() +
  scale_y_log10() +
  theme_classic(base_size = 16) +
  ggtitle("a_l2 = 0.1")
  
##################### ##################### #####################
  p0_22 <- scm_base_parameters("FF16")
  p0_22$strategy_default$a_l1 <- 2.17
  p0_22$strategy_default$a_l2 <- 0.5
  p0_22$strategy_default$recruitment_decay <- 5
  
  control = scm_base_control()
  control$ode_tol_rel <- 1e-4
  control$ode_tol_abs <- 1e-4
  control$schedule_eps <- 0.0005
  
  patch22 <- run_mypatch_edit(p0 = p0_22, ctrl = control) %>%
    tidy_patch() %>%
    FF16_expand_state()
  
  patch22_total <- 
    patch22$species %>% 
    integrate_over_size_distribution() %>%
    mutate(
        diameter_stem = (2 * sqrt (area_stem / pi)),
        diameter_stem_av = diameter_stem/density,
        across(c(diameter_stem, area_stem, area_leaf, height, mass_above_ground), 
        ~.x/density, .names = "{.col}_av")
    )  
  
  p22 = patch22_total %>%
    filter(time > 2) %>%
    ggplot(aes(diameter_stem_av, density)) + 
    geom_line() + 
    scale_x_log10() +
    scale_y_log10() +
    theme_classic(base_size = 16) +
    ggtitle("a_l2 = 0.5")
  
##################### ##################### #####################
  p0_33 <- scm_base_parameters("FF16")
  p0_33$strategy_default$a_l1 <- 2.17
  p0_33$strategy_default$a_l2 <- 1
  p0_33$strategy_default$recruitment_decay <- 5
  
  control = scm_base_control()
  control$ode_tol_rel <- 1e-4
  control$ode_tol_abs <- 1e-4
  control$schedule_eps <- 0.0005
  
  patch33 <- run_mypatch_edit(p0 = p0_33, ctrl = control) %>%
    tidy_patch() %>%
    FF16_expand_state()
  
  patch33_total <- 
    patch33$species %>% 
    integrate_over_size_distribution() %>%
    mutate(
        diameter_stem = (2 * sqrt (area_stem / pi)),
        diameter_stem_av = diameter_stem/density,
        across(c(diameter_stem, area_stem, area_leaf, height, mass_above_ground), 
        ~.x/density, .names = "{.col}_av")
    ) 

p33 = patch33_total %>%
    filter(time > 2) %>%
    ggplot(aes(diameter_stem_av, density)) + 
    geom_line() + 
    scale_x_log10() +
    scale_y_log10() +
    theme_classic(base_size = 16) +
    ggtitle("a_l2 = 1")
```

## Checking values with birth rate
### 100, 200, 300, 600
### STLs are not affected by birth
```{r}
birth_rate_runs <- c(100, 200, 300, 600) %>% set_names(.) %>% map(~grow_tidy_expand_totals(birth_rate = .x))
## why no difference? where is birthrate value to check?

birth_rate_total <-
  birth_rate_runs %>% map_dfr(~.x$patch_total, .id = "run") %>%
  mutate(run = recode(run, '1' = 'birth rate = 100', '2' = 'birth rate = 200', '3' = 'birth rate = 300', '4' = 'birth rate = 600'))

thinning_plot(birth_rate_total, xvar = "diameter_stem_av", xlims = c(1e-04, 1e-01), ylims = c(1e-01, 50), ylab = "Stem Diameter (m)", title = "Self-thinning under varied birth rates")
#something up with birth_rate? affecting seed decay?
```

## Checking values with k_l
### 0.3, 1, 2, 2.7
```{r}
kl_runs <- c(0.3, 1, 2, 2.7) %>% set_names(.) %>% map(~grow_tidy_expand_totals(k_l = .x))

kl_total <-
  kl_runs %>% map_dfr(~.x$patch_total, .id = "run") %>%
  mutate(run = recode(run, '1' = 'k_l = 0.2', '2' = 'k_l = 0.4565855', '3' = 'k_l = 0.7', '4' = 'k_l = 0.9'))
```

## Checking values with b_lf1
### 0.8273474, 3, 5, 7
```{r}
blf1_runs <- c(0.6, 0.8273474, 0.95) %>% set_names(.) %>% map(~grow_tidy_expand_totals(B_lf1 = .x))

blf1_total <-
  blf1_runs %>% map_dfr(~.x$patch_total, .id = "run") %>%
  mutate(run = recode(run, '1' = 'b_lf1 = 0.6', '2' = 'b_lf1 = 0.8273474', '3' = 'b_lf1 = 0.95'))
```

## Checking values with b_lf2
### 0.25, 0.5, 0.8
```{r}
blf2_runs <- c(0.25, 0.5, 0.8) %>% set_names(.) %>% map(~grow_tidy_expand_totals(B_lf2 = .x))

blf2_total <-
  blf2_runs %>% map_dfr(~.x$patch_total, .id = "run") %>%
  mutate(run = recode(run, '1' = 'b_lf2 = 0.25', '2' = 'b_lf2 = 0.5', '3' = 'b_lf2 = 0.8'))
```

## Checking values with d_I
### 0, 0.05, 0.1, 0.2
### higher DI mortality leads to greater average stem diameter, 
### density roughly same
```{r}
dI_runs <- c(0.01, 0.05, 0.1, 0.2) %>% set_names(.) %>% map(~grow_tidy_expand_totals(d_I = .x))

dI_total <-
  dI_runs %>% map_dfr(~.x$patch_total, .id = "run") %>%
  mutate(run = recode(run, '1' = 'd_I = 0', '2' = 'd_I = 0.05', '3' = 'd_I = 0.1', '4' = 'd_I = 0.2'))

thinning_plot(dI_total, xvar = "diameter_stem_av", xlims = c(-0.02, 0.1), title = "Self-thinning under varied DI mortality")

thinning_plot(birth_rate_total, xvar = "diameter_stem_av", xlims = c(0.02, 1.0), ylab = "Stem Diameter (m)", title = "Self-thinning under varied birth rates")
```


# We need to check selected trait values against empirical data
## Woody plant trait values from Austraits database
```{r}

austraits <- read.csv("../self-thinning/austraits-3.0.2/traits.csv", header = TRUE)
woodytraits <- read.csv("../self-thinning/austraits-3.0.2/woodiness_2022.03.19.csv", header = TRUE)

sort(unique(austraits$trait_name)) # what traits are recorded?

austraits <- woodytraits %>% select(taxon_name, woodiness_recoded) %>% right_join(austraits) %>% mutate(ID = row_number())

wood_density_vals <- austraits %>% filter(trait_name == "wood_density", woodiness_recoded == "woody") %>% mutate(ID = row_number())
leaf_N <- austraits %>% filter(trait_name == "leaf_N_per_area", woodiness_recoded == "woody") %>% mutate(ID = row_number())
LMA_vals <- austraits %>% filter(trait_name == "???", woodiness_recoded == "woody")
seed_mass_vals <- austraits %>% filter(trait_name == "seed_mass", woodiness_recoded == "woody", value < 100) %>% mutate(ID = row_number())
## all_traits <- austraits %>% filter(trait_name == "wood_density" | trait_name == "leaf_N_per_area" | trait_name == "seed_mass", woodiness_recoded == "woody") %>% mutate(ID = row_number())

## filter out woody stuff

max(leaf_N$value)
min(leaf_N$value)

min(seed_mass$value)
max(seed_mass$value)

# plot range of trait values:

ggplot(LMA_vals, aes(as.numeric(value))) + geom_histogram(binwidth = 0.25) + xlim(0, 25) + xlab("LMA g/m^2")
ggplot(leaf_N, aes(as.numeric(value))) + geom_histogram(binwidth = 0.25) + xlim(0, 25) + xlab("N per unit leaf area g/m^2")
ggplot(wood_density_vals, aes(as.numeric(value))) + geom_histogram(binwidth = 0.01) + xlab("Wood Density")
ggplot(seed_mass_vals, aes(as.numeric(value))) + geom_histogram(binwidth=0.5) + xlab("Seed Mass")
```

## BAAD data
```{r}
baad_data <- read.csv("../self-thinning/baad_data.csv", header = TRUE)
baad_data_dictionary <- read.csv("../self-thinning/baad_dictionary.csv", header = TRUE) %>% rbind()
baad_data_dictionary$label

## "light environment" = light, "age" = age, "leaf area" = a.lf, "height" = h.t, "stem area at base/breast height/crown base" = a.stba/a.stbh/a.stbc, "dbh" = d.bh, "leaf mass" = m.lf, "total mass" = m.to, "wood density" = r.st, "leaf mass per area" = ma.ilf, "aboveground mass" = m.so, "leaf [nitrogen]" = n.lf
```

# age vs LA
```{r}
# age vs LA values...

baad_data %>%
  ggplot(aes(age, a.lf, col = studyName)) +
  geom_point(show.legend = FALSE) +
  theme_classic()

baad_data %>%
  ggplot(aes(age, a.lf, col = studyName)) +
  geom_point(show.legend = FALSE) +
  geom_point(data = , )
  theme_classic() +
  xlim(0, 300) +
  ylim(0, 800)
```

# DBH vs age
```{r}
baad_data %>%
  ggplot(aes(age, a.lf, col = studyName)) +
  geom_point(show.legend = FALSE) +
  theme_classic()

baad_data %>%
  ggplot(aes(age, a.lf, col = studyName)) +
  geom_point(show.legend = FALSE) +
  geom_point(data = , )
  theme_classic() +
  xlim(0, 300) +
  ylim(0, 800)
```

# DBH vs LA
```{r}
baad_data %>%
  ggplot(aes(age, a.lf, col = studyName)) +
  geom_point(show.legend = FALSE) +
  theme_classic()

baad_data %>%
  ggplot(aes(age, a.lf, col = studyName)) +
  geom_point(show.legend = FALSE) +
  geom_point(data = , )
  theme_classic() +
  xlim(0, 300) +
  ylim(0, 800)
```
