---
title: "Plotting empirical data against plant"
output: html_document
date: '2022-08-24'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results='hide')
```

```{r, eval = FALSE}
remotes::install_github("traitecoevo/plant", branch = "develop", force = TRUE)
```

```{r load libraries}
library(plant)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(ggrepel)
library(scatterplot3d)
source("R/grow_tidy_expand_totals.R")
source("R/run_mypatch.R")
source("R/thinning_plot.R")
source("R/plot_stand.R")
source("R/find_closest_timestamps.R")
```

```{r activate logger}
plant_log_console()
```



# Here we compare model output with empirical datasets
First, grow a stand - traits and other preset values not altered
Initial seed rain is 200, a_l1 = 2.17 (height of plant with LA 1m^2), a_l2 = 0.5 (exponent of relationship between height and LA), lma = 0.1978791, rho = 608, hmat = 16.59587, B_lf1 = 0.8273474, B_lf2 = 0.5, recruitment decay = 5
```{r}
examplestand <- grow_tidy_expand_totals()
stand_totals <- examplestand$patch_total  
```

# Generate self-thinning lines from slope and intercept values obtained through statistical models of past papers

1. Trouve et al stand data with one species' slope and intercept values obtained from 4 different statistical models
Intercept values are logged so take exponential
```{r}
fitted_trouve <- empirical_studies %>% filter(study == "trouve_et_al") %>%
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% # use max/min values for this
  mutate(density = intercept*diameter_stem_av^as.numeric(slope),
         study = "Trouve_et_al",
         slope = as.numeric(slope)) %>%
  select(-min_dq, -max_dq)

```
start/end points of empirical data (different stats methods?)

"Of the two self-thinning methods applied to the Dq-Nallometry, SFA seemed to provide better upperNboundary estimates than OLS, but sometimesfailed as it was more demanding in terms of data(e.g., with E. camaldulensis). Of the three mortal-ity models, self-thinning estimates were more con-sistent and error distribution assumptions betterbehaved for NB GLM than for Poisson GLM andNLS."

E. camaldulendis: 
Dq (cm): 6-59, 9-26, 6-53,  
N (tree/ha): 112-5513, 1195-5513, 171-5513 

E. delegatensis: 
Dq (cm): 7-67, 9-63, 7-64, 9-40,  
N (tree/ha): 201/272/201/420 - 7259 

E. nitens: 
Dq (cm): 14-50, 18-50, 14-48,  
N (tree/ha): 224-1145, 224-1156, 225-1445

E. obliqua: 
Dq (cm): 6-62, 8-62, 6-60,  
N (tree/ha): 99-11475, 224-11100, 123-11475

E. regnans: 
Dq (cm): 3-114, 3-112, 3-111, 13-92,  
N (tree/ha): 27-39288, 64-39288, 27-39288, 115-1225

E. sieberi: 
Dq (cm): 3-80, 8-47, 3-79, 9-35,  
N (tree/ha): 49-14247, 478-9525, 49-14247, 364-4169


2. Condes et al stand data with species slope and intercept values obtained for 2 species (pine, beech) in 6 locations (12 sets of slope values)
Take natural log of intercepts
Max stand denstiy index i.e. max # stems per hectare for reference QMD of 25cm
```{r}
fitted_condes <-
  tibble(
  slope = c(-1.545, -1.539, -2.226, -1.494, -1.592, -1.704, -1.918, -1.469, -2.244, -1.850, -1.831, -1.720),
  intercept = c(190722, 154611, 1213554, 127071, 185631, 210150, 577162, 110800, 1615685, 341925, 428430, 254156), 
  species = c(rep(1, 6), rep(2, 6)),
  replicate = c(1, 2, 3, 4, 5, 6, 1, 2, 3, 4, 5, 6)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Condes_et_al")
```
start/end points of empirical data
pine: 
N (tree/ha): 247-3656, 144-4317, 80-5363, 59-3533, 42-3836, 80-4000,  
Dq (cm): 
11.2 37.6 
10.1 50.1
10.1 52.9 
10.0 49.6 
10.1 54.9 
10.0 54.9

beech: 
N (tree/ha): 70-4730, 242-2913, 100-3623, 61-4838, 57-6057, 75-2600
Dq (cm): 
10.2-54.7
10.4-44.7
11.4-49.7
10.1-54.7
10.1-54.9
10.2-53.8


3. Weiskittel, Gould & Temesgen stand data
Species slope and intercept values obtained for 3 species (Douglas-fir, Red alder, Western hemlock) using OLS & SFA methods
Density: trees per ha, mean tree size: QMD -- both log transformed
```{r}
fitted_wgt <-
  tibble(
  slope = c(-1.2618, -1.5473, -1.1133, -1.2208, -1.6545, -1.7336), #OlS, SFA
  intercept = c(10.7509, 11.2087, 9.7315, 10.8665, 12.1299, 13.0391) %>% exp(), #OlS, SFA
  species = c(1, 1, 2, 2, 3, 3),
  replicate = c(1, 2, 1, 2, 1, 2)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Weiskittel_Gould_Temesgen")
```
Douglas-fir
N: 98.8 - 4694.9 (stems/ha)
Dq: 4.6 - 73.6 (cm)

Red alder
N: 3660.5 - 37.1 
Dq: 0.7 - 48.8

Western hemlock
N: 98.8 - 5782.1
Dq: 7.7 - 43.7


4. Rivoire and Le Moguedec
Oak, beech and spruce modelled with stochastic frontier estimation & generalised
optimisation (-spruce intercept n.a.) (5 STLs)
```{r}
fitted_riv_mog <-
  tibble(
  slope = c(-1.77, -1.76, -1.60, -1.86, -1.59),
  intercept = c(12.22, 11.69, 11.78, 12.17, 11.99) %>% exp(),
  species = c(1, 1, 2, 2, 3),
  replicate = c(1, 2, 1, 2, 1)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Rivoire_Le_Moguedec")
```
start/end points of empirical data
Not given??


5. Gosper Euc data: E. salubris y = 4.15 - 1.33x
Density per ha, QMD at 10cm above base ???
```{r}
fitted_gosper <- 
  tibble(
  slope = -1.33,
  intercept = 4.15 %>% exp(), species = seq_len(1)
  ) %>%
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Gosper")
```
start/end points of empirical data


Hammond & Niklas
A. alba - uses mass not stem diameter...
--

6. Eljatib & Weiskittel
```{r}
fitted_elj_weis <-
  tibble(
  slope = c(-1.2741, -1.4742, -1.33, -2.2396, -2.7194),
  intercept = c(11.4787, 12.257, 11.71, 14.535, 16.352) %>% exp(),
  species = 1,
  replicate = c(1, 2, 3, 4, 5)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Eljatib_Weiskittel")
```
can't find...


7. Pretzsch & Rio 2020
N. spruce, S. pine, D-fir, E. beech, s. oak
Central Europe long-term experimental plots
```{r}
fitted_pret_rio <-
  tibble(
  slope = c(-1.650, -1.789, -1.534, -1.887, -1.753),
  intercept = c(12.560, 12.473, 11.811, 12.865, 12.254) %>% exp(),
  species = seq_len(5)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Pretzsch_Rio")
```
not given...


8. Andrews et. al. 2018
Forests in Acadian Region, Northeastern North America. 15 species of soft and hardwood
```{r}
fitted_andrews <-
  tibble(
  slope = c(-1.96, -1.32, -1.63, -1.57, -1.28, -1.45, -1.38, -1.66, -1.68, -1.77, -1.62, -1.84, -1.39, -1.47, -1.33),
  intercept = c(12.82, 10.98, 11.82, 11.78, 10.98, 11.61, 11.38, 11.88, 11.89, 12.09, 11.70, 12.39, 11.18, 11.57, 10.97) %>% exp(),
  species = seq_len(15)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Andrews_et_al")
```
start/end points of empirical data (stems/ha, cm)
Balsam Fir 55.00 24975.00, Dq: 4.48 29.57
White Spruce: N: 74.35 - 12427.95, Dq: 4.49 36.87
Black Spruce: N: 75.00 - 24225.75, Dq: 4.49 30.25
Red Spruce: N: 59.48 - 18125.00, Dq: 4.53 36.94
E. White Pine: N: 59.48 - 23391.00, Dq: 4.59 46.78
N. White Cedar: N: 64 - 100.00 10879.22, Dq: 4.49 40.57
E. Hemlock: N: 59.48 - 16512.10, Dq: 4.50 46.77
Red Maple: N: 59.48 - 13026.49, Dq: 4.50 50.45
Sugar Maple: N: 56.83 - 9010.63, Dq: 4.53 48.89
Yellow Birch: N: 56.83 - 10641.28, Dq: 4.69 44.16
Paper Birch: N: 75.00 - 19480.50, Dq: 4.56 35.47
American Beech: N: 56.83 - 8254.79, Dq: 4.50 42.57
White Ash: N: 59.31 - 7670.60, Dq: 4.50 46.93
Bigtooth Aspen: N: 74.35 - 20905.75, Dq: 4.52 31.27
N. Red Oak: N: 74.35 - 7713.93, Dq: 4.99 51.83 


9. Aguirre, A., del Río, M., & Condés, S. 2018
Pine species: P. halepensis Mill., P. nigra Arn., P. pinea L., P. pinaster Ait., and P. sylvestris L.
(eqn 1: log(N-max) no climate, eqn 2b: log(N-max) with climate)
```{r}
fitted_ag_rio_cond <-
  tibble(
  slope = c(-1.7262, -1.7950, -2.1220, -2.2647, -1.8292, -1.7009, -1.7943, -1.8759, -1.9825, -1.7298),
  intercept = c(12.607, 12.619, 13.385, 13.426, 12.317, 7.982, 12.642, 11.819, 13.341, 9.784) %>% exp(),
  species = c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5), 
  replicate = c(1, 2, 1, 2, 1, 2, 1, 2, 1, 2)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Aguirre_Río_Condés")
```
start/end points of empirical data (stems/ha, cm)
P. sylvestris
78 - 4333, Dq: 
10.0 49.6
P. pinea
29 - 3583, Dq: 
10.0 49.7
P. halepensis
N: 14 - 2451, Dq:	
10.0 48.9
P. nigra
N: 29 - 4623,	Dq: 
10.0 49.7
P. pinaster
N: 50 - 4520,	Dq: 
10.0 49.9


10. Ge, F. et al 2017
China fir (Cunninghamia lanceolata) pure stand plantations, !! varied site quality, mixed species
```{r}
fitted_ge <-
  tibble(
  slope = c(-4.298, -3.417, -2.865, -1.590, -1.694),
  intercept = c(18.933, 16.950, 15.695, 12.365, 12.544) %>% exp(),
  species = seq_len(5)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Ge_et_al")
```
start/end points of empirical data
N (stems/ha): 104 - 6791
Dq (sm): 5.7 - 25.0


Stankova, T. V., & Diéguez-Aranda, U. 2017. Galicia (NW Spain) and Scots pine plantations in Bulgaria.
Even-aged stands of four tree species
Radiata pine (Pinus radiata), Scots pine (Pinus sylvestris), downy birch (Betula pubescens), English oak (Quercus robur)
Use mean stem volume not diameter...
```{r}
#fitted_stan_dieg <-
#  tibble(
#  slope = c(-1.26, -1.03, -1.16, -0.71),
#  intercept = c(3621, 6169, 11279, 2321),
#  species = seq_len(4)
#  ) %>% 
#  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
#  mutate(density = intercept*diameter_stem_av^slope,
#         study = "Stankova_Diéguez-Aranda")
```


Brunet-Navarro, P., Sterck, F. J., Vayreda, J., Martinez-Vilalta, J., & Mohren, G. M. J. (2016).
The minimum tree density threshold where self-thinning take place was 732 trees ha−1 for P. halepensis, 600 trees ha−1 for P. nigra, 579 trees ha−1 for P. sylvestris and 581 trees ha−1 for 
ALSO: slopes with varied climate!! Add these
```{r}
fitted_brunet <-
  tibble(
  slope = c(-0.533, -0.540, -0.589, -0.571),
  intercept = c(14.85, 16.01, 18.31, 18.54) %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Brunet-Navarro_et_al")
```
annoying data...
see: https://annforsci.biomedcentral.com/articles/10.1007/s13595-016-0585-y/tables/1


Vospernik, S., & Sterba, H. 2015
Norway spruce, European beech, Scots pine, European larch, oak, Austrian pine, silver fir, ash, stone pine and hornbeam. Sycamore, grey alder, birch, black locust and black alder are found on less than 100 plots.
```{r}
fitted_vospernik_sterba <-
  tibble(
  slope = c(-1.753, -1.633, -1.936, -1.750, -1.845, -1.500, -1.941, -1.800, -2.000, -1.939),
  intercept = c(12.86, 12.41, 13.00, 12.77, 13.20, 11.62, 13.25, 12.62, 13.33, 13.22) %>% exp(), # only 0.5 quartile for birch and grey alder
  species = seq_len(10)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Vospernik_Sterba")
```
Norway spruce	
5.0	81.5
8	10,185	

Silver fir	
6.5	84.8
15	3608

European larch	
5.3	81.1
8	4427

Scots pine	
5.1	49.8
21	6741

Austrian pine
11.2	47.8
57	5021

Stone pine	
6.5	67.4
18	1266

European beech	
5.0	85.5
9	6224

Common oak
5.1	79.5
9	3663

Hornbeam
5.3	48.1
22	8494

Ash
5.1	66.2
12	6478

Birch
5.1	52.2
19	5511

Grey alder
5.7	34.7
42	5505

Sycamore
5.3	76.6
9	3103

Black locust
5.2	24.6
284	5238

Black alder
5.0	32.4
49	4749


Wang, N., Gao, J., Xu, S.-S., Zhang, W.-P., & Wang, G.-X. 2013
Mixed montane, evergreen montane, deciduous montane, shrub. Mixed species.

humid area: mean α = -1.74, predicted α =  -1.58
semi-arid area: mean α = -1.27, predicted α =  -1.29
arid area: mean α = -0.95, predicted α = -0.94

intercepts??
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

11. Charru, M., Seynave, I., Morneau, F., Rivoire, M., & Bontemps, J.-D. 2012
Pure and even-aged plots of 11 broadleaved and coniferous species
```{r}
fitted_charru_et_al <-
  tibble(
  slope = c(-1.809, -1.758, -1.911, -1.941, -1.615, -1.881, -1.653, -1.711, -1.878, -1.779, -1.646),
  intercept = c(12.270, 12.138, 12.681, 12.950, 11.993, 12.512, 12.104, 11.982, 13.086, 12.621, 12.133) %>% exp(),
  species = seq_len(11)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "Charru_et_al")
```

Pubescent oak
N: 39.3	1,984.5
Dq: 15.0	36.5

Pedunculate oak	
N: 39.3	2,205.0	
Dq: 15.1	36.9

Sessile oak	
N: 39.3	2,544.5	
Dq: 15.0	30.0

Common beech	
N: 14.1	2,710.3
Dq: 16.0	40.0

Scots pine	
N: 39.3	3,301.0	
Dq: 15.0	33.0

Aleppo pine	
N: 28.3	2,672.2	
Dq: 15.0	40.0

Corsican pine	
N: 67.6	1,945.2	
Dq: 18.0	49.8

Maritime pine	
N: 39.3	3,517.1	
Dq: 16.0	37.0

Norway spruce	
N: 42.4	3,652.3	
Dq: 15.0	45.0

Silver fir	
N: 70.7	2,908.0	
Dq: 15.0	44.9

Douglas fir
N: 28.3	2,661.2	
Dq: 20.0	49.6

Tang, S., Meng, F.-R., & Meng, C. H. 1995
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

Forrester et al 2021
Data from temperate Australian Eucalyptus plantations (436 plots), subtropical forests in China (88 plots), and temperate forests in Switzerland (1055 plots)
https://www.sciencedirect.com/science/article/pii/S0378112721000256?via%3Dihub
predicted self-thinning trajectories (results not shown)???
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```
Abies alba	(15/2779) (10.8/62.7)
Eucalyptus globulus	(131/1718)(6.2/46.9)
Eucalyptus nitens	(100/1735)(13.1/47.5)
Fagus sylvatica	(10/1664)(11.5/52.1)
Fraxinus excelsior	 (8/946)(11.4/55.7)
Larix decidua	(8/1769)(11.8/70.8)
Picea abies	(19/3840)	(10.6/57.3)
Pinus strobus	(29/1988)(15.5/61.7)
Pinus sylvestris	(10/4814)(11.8/68.6)
Pseudotsuga menziesii	(8/2500) (14.4/66.9)	
Quercus petraea or Q. robur	(3/1700)(9/75.4)
Subtropical China (507/3119) (5.5/22)

Morris, Morris & Myerscough data: Ocimum basilicum (herb?)
```{r}
fitted_other <-
  tibble(
  slope = c(),
  intercept = c() %>% exp(),
  species = seq_len(4)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = )
```

Mixing effects on growth efficiency in mixed pine forests 
José Riofrío, Miren del Río, Felipe Bravo
```{r}
fitted_riof_rio_brav <-
  tibble(
  slope = c(-1.789, -1.929),
  intercept = c(12.801, 13.218) %>% exp(),
  species = seq_len(2)
  ) %>% 
  expand_grid(diameter_stem_av = 10^seq(0,2)) %>% 
  mutate(density = intercept*diameter_stem_av^slope,
         study = "riofrío_del_río_bravo")

```
start/end points of empirical data (trees/ha, cm)
Mixed plots:
Scots pine natural
N: 14.2 - 1322.7 
Dq: 9.8 - 50.7
Maritime pine natural
N: 19.2 - 1177.7 
Dq: 10.3 - 41.6
Scots pine plantation
N:  88.4 - 1987.6 
Dq: 8.8 - 35.8
Maritime pine plantation
N: 	28.3 - 1481.9
Dq: 11.2 - 31.7

N
Total natural:87.1 - 1479.9 	   		
Total plantation: 226.4 - 3087.5
Dq
Total natural: 11.7 - 39.5	                	          
Total plantation: 10.7 - 33.6

Pure plots:
Scots pine natural
N: 65.1 - 2893.0 
Dq: 10.5 - 51.5
Maritime pine natural
N: 104.1 - 2291.8
Dq: 10.0 - 46.1
Scots pine plantation
N: 240.5  - 3755.9 
Dq: 8.9 - 32.5
Maritime pine plantation
N: 283.0 - 1955.8
Dq: 10.7 - 33.5

# Combine studies:
```{r}
prev_studies <- as_tibble(bind_rows(fitted_condes, fitted_gosper, fitted_trouve, fitted_riv_mog, fitted_wgt, fitted_elj_weis, fitted_ag_rio_cond, fitted_andrews, fitted_charru_et_al, fitted_ge, fitted_pret_rio, fitted_brunet, fitted_vospernik_sterba)) %>%
  mutate(replicate = replace_na(replicate, 1))
```

# Typical self thinning line using plant:
```{r}
examplestand <- grow_tidy_expand_totals(a_dG2 = 30)

# correct for units -- m to cm, m^2 to ha
patch_totals <- examplestand$patch_total %>%
  mutate(
    diameter_stem_av = (diameter_stem_av*100),
    density=density*10000,
    study = "plant",
    replicate = 1
    ) %>%
  group_by(study, species)

  ggplot(patch_totals, aes(diameter_stem_av, density)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic(base_size = 18)
  
###  
examplestand <- grow_tidy_expand_totals(hmat = 15, a_dG2 = 30)
patch_totals <- examplestand$patch_total %>%
  mutate(
    study = "plant",
    replicate = 1,
    diameter_stem_av = (diameter_stem_av*100),
    density=density*10000,
    ) %>%
  group_by(study, species) %>%
  filter(density>0.001)

# plot STL
patch_totals %>%
  ggplot(aes(diameter_stem_av, density)) +
  geom_line() +
  scale_x_log10(name = "mean stem diameter (cm)") +
  scale_y_log10(name = "density (ha)", labels = scales::comma) +
  theme_classic(base_size = 18)

ggsave(typical_stand, filename = "../self-thinning/typical_stand.pdf", width = 6, height = 5)
```

# Plot self-thinning lines from compilation of datasets of major studies plotted with modelled STL using plant
```{r}
plant_and_empirical = ggplot(mapping = aes(diameter_stem_av, density, group = paste(study, species, replicate))) +
  geom_line(data = prev_studies, col = "grey") +
  geom_line(data = patch_totals, size = 1, col = "black") +
  geom_abline(intercept = 5.3, slope = -3.8/2, linetype = "dashed") +
  scale_x_log10(limits = c(0.1, 500), labels = scales::comma) +
  scale_y_log10(labels = scales::comma) +
  theme_classic(base_size = 18) +
  xlab("log mean stem diameter (cm)") +
  ylab("log density (ha)")


ggsave(plant_and_empirical, filename = "../self-thinning/plant_and_empirical.jpeg", width = 6, height = 5)
plot_size_distribution(examplestand$patch_expand$species)

# x and y limits from each study
```









?? fitted_brunet 
intercept = c(14.85, 16.01, 18.31, 18.54) %>% exp()

Brunet-Navarro, P., Sterck, F. J., Vayreda, J., Martinez-Vilalta, J., & Mohren, G. M. J. (2016). Self-thinning in four pine species: An evaluation of potential climate impacts. Annals of Forest Science, 73(4), 1025–1034. https://doi.org/10.1007/s13595-016-0585-y

?? fitted_gosper
intercept = 4.15 %>% exp()

Gosper, C. R., Yates, C. J., Cook, G. D., Harvey, J. M., Liedloff, A. C., McCaw, W. L., Thiele, K. R., & Prober, S. M. (2018). A conceptual model of vegetation dynamics for the unique obligate-seeder eucalypt woodlands of south-western Australia. Austral Ecology, 43(6), 681–695. https://doi.org/10.1111/aec.12613
